The Creation of a Programming Language

### 2022-09-03 paper.tex
paper make target, continue writing paper, add bibliography

\documentclass[12pt,a4paper,man]{apa7}
\usepackage{listings}

\author{Florian Donnelly}
\title{The Creation of a Programming Language}
\date{August 2022}
\affiliation{Gymnasium Burgdorf}
\shorttitle{Creation of a Programming Language}
\abstract{
    This is an abstract...
}

% Defining custom commands
\newcommand{\code}[3] {
    \lstset{language=C,numbers=left,basicstyle=\scriptsize,
        frame=single,caption=src/#1,keywordstyle=\color{blue},
        stringstyle=\color{green},commentstyle=\color{gray},
        morecomment=[l][\color{magenta}]{\#}}
    \lstinputlisting{src/#1}
}

\begin{document}
\maketitle

\tableofcontents\newpage

\section{Introduction}

\code{main.c}

\section{Theory}
\section{Implementation}
\section{Results}
\section{Discussion}
\section{Index}
\section{Closing Words}
\section{Declaration of Autonomy}
\section{Appendix}

>>>>

\documentclass[12pt,a4paper,man]{apa7}
\usepackage{listings}
\usepackage{setspace}
\usepackage{times} % Use Times New Roman font
\usepackage[style=apa]{biblatex}

\addbibresource{bibliography.bib}

\author{Florian Donnelly}
\title{The Creation of a Programming Language}
\authorsaffiliations{Gymnasium Burgdorf}
\shorttitle{Creation of a Programming Language}
\abstract{
    This is an abstract...
}

% Defining custom commands

% include C code from file #1 from lines #2 to #3
\newcommand{\code}[3] {
    \lstset{language=C,numbers=left,basicstyle=\scriptsize,
        frame=single,caption=src/#1,keywordstyle=\color{blue},
        stringstyle=\color{green},commentstyle=\color{gray},
        morecomment=[l][\color{magenta}]{\#}}
    \lstinputlisting[firstnumber=#2,firstline=#2,lastline=#3]{src/#1}
}

\begin{document}

\onehalfspacing
\maketitle\tableofcontents\newpage

% the name of the programming language
\newcommand{\name}{-the language- }

\section{Introduction}
\subsection{Relevance of the Topic}
\subsection{Goals}

\code{main.c}{2}{6}

\section{Theory}
\subsection{What this is and what this is not}
The final product in the process of creating this programming language will 
not be perfect, or in fact, be used or adopted by anyone at all, and it is not
intended to be. Creating a good, or even usable, programming language 
implementation by today's  standards means to create an entire ecosystem of 
tools, not only including a compiler or interpreter, but also a package manager
for dealing with dependencies and a platform for people to share their work.

The process used in this paper is more so a very core implementation of what
a programming language, only containing the very necessary features.

\section{Implementation}
\subsection{Used Software}
At the heart of creating a program in C stands the C compiler. This project uses
\emph{gcc}, the GNU Compiler Collection, which, as the name suggests, also 
supports different languages than C. Substantiating the development process is 
\emph{GNU Make}, a general-purpose build system. It handles tasks like building
the executable step-by-step from the project source code and directory clean-up.
The C ecosystem is very easy to use and fundamentally supported by 
\emph{GNU+Linux}, the operating system used to program on.
The main code editor used was \emph{Microsoft Visual Studio Code}, providing great
features like built-in git integration.
\emph{Git} is the project management and tracking software used to back-up and share
the project files on \emph{GitHub.com}, a Microsoft hosted git repository server to
store and collaborate on source code.

\section{Results}
\section{Discussion}

\subsection{Interesting Ideas untouched}
Creating the ideal programming language takes a lot of planning and sketching
out. This section of the paper gives insight on interesting ideas or concepts
that are used in other programming languages, not \name.

\subsubsection{Foreign Function Interfaces}
A foreign function interface\footcite{FFI} often abbreviated as FFI,
is a mechanism by which one programming language can call functions written 
in another. This makes a language more attractive, as it can be used with
existing libraries from another language, instead of functions having to be
re-written. Enabling C function calls often allows a certain language to
directly interact with system libraries or the operating system itsself.

Foreign function interfaces are available in many modern languages.
Examples are Java, Python and Rust.
Java enables users to call C, C++ and even assembly code with the Java Native
Interface\footcite{JNI}, JNI for short, whereas within Python C function can
be run with the standard ctypes\footcite{ctypes} library.

Foreign function interfaces are specially difficult to implement, or implement
right. Let us assume a programming language interpreter \name written in C that
would like to call some functions that are part of yet another C program or
library. The problem lies in the fact that at runtime, C does not know
structure of the functions to be called, say return- or argument-types, which
is by the way where the term foreign functions originates from. This is why the
\name developer would have to provide that information by re-declaring the
function inside \name, so that, with some trickery, data can be passed to and
received from the function in the correct format, preventing a segmentation
fault. The libffi\footcite{libffi} C-library, which is also used by Java and
Python for this purpose, can be used to load and call such foreign functions.

\subsubsection{Reflection}
Reflection\footcite{reflection}, also called reflective programming, describes
the ability of a program to examine and modify its own structure and behaviour 
at runtime. An example would be self-modifying code. This can easily be
achieved in assembly language, which inherently does not differenciate between
data and instructions.
Another example is Java, providing methods and classes in the java.lang.reflect
package that enable developers to examine and change properties or functions
of objects.
The advantage of reflection over not having it is that it allows for non-static
programs and program flows, effectively making smaller executables and more
resource efficient programs, as those programs can adjust themselfs to run
differently under different conditions at runtime.
The easiest and simplest way to go about implementing Reflection in \name is
certainly to allow the interpretation of strings at runtime. That way, strings
holding program instructions can be created and modified and later executed.

\section{Index}
\section{Closing Words}
\section{Declaration of Autonomy}
Hereby I, Florian Donnelly, declare to have written this paper, also including
the entire source code of \name by myself using resources listed under the
references section.

\section{Appendix}

\newpage\printbibliography[heading=bibintoc]
